---
title: "EDA CH4"
output: html_document
date: "2025-03-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(reshape2)
```
Objective:
To analyze the relationship between global CO₂ concentrations and global temperature anomalies from 2000 to 2010, identifying potential correlations and trends. The study aims to determine whether rising CO₂ levels correspond to an increase in global temperatures, providing insights into climate change patterns.

Hypothesis:
Null Hypothesis (H₀): There is no significant correlation between global CO₂ concentrations and global temperature anomalies from 2000 to 2010.
Alternative Hypothesis (H₁): There is a significant positive correlation between global CO₂ concentrations and global temperature anomalies from 2000 to 2010, indicating that rising CO₂ levels contribute to temperature increases.
```{r}
df_anomaly = read.csv("D:\\DS-B3\\Spatial\\Analyzing-Global-Temperature-and-CO-Emission-Trends-2000-2010-\\dataset\\anomaly_2000_2010.csv")
df_ch4 = read.csv("D:\\DS-B3\\Spatial\\Analyzing-Global-Temperature-and-CO-Emission-Trends-2000-2010-\\dataset\\ch4_2000_2010.csv")
df_co2 = read.csv("D:\\DS-B3\\Spatial\\Analyzing-Global-Temperature-and-CO-Emission-Trends-2000-2010-\\dataset\\co2_2000_2010.csv")
```

```{r}
head(df_anomaly)
head(df_ch4)
```

```{r}
colnames(df_ch4)
```
EDA:
Step 1: Understand the Problem and the Data
Step 2: Import and Inspect the Data
Step 3: Handling Missing Values
Step 4: Explore Data Characteristics
Step 5: Perform Data Transformation
Step 6: Visualize Data Relationships
Step 7: Handling Outliers
Step 8: Communicate Findings and Insights
```{r}
names(df_ch4)[names(df_ch4) == "average"] = "average_ch4"
names(df_ch4)[names(df_ch4) == "trend"] = "trend_ch4"
```
Topic:
Investigating the Impact of CO₂ and CH₄ Emissions on Global Temperature Anomalies (2000-2010)

Objective:
This study aims to explore the relationship between global temperature anomalies and greenhouse gas emissions (CO₂ and CH₄) over the period 2000-2010. By analyzing monthly trends, we seek to determine whether increases in atmospheric CO₂ and CH₄ concentrations correspond to significant temperature changes. The study will use statistical correlation and trend analysis to assess the strength and direction of these relationships.

Hypotheses:
H₀ (Null Hypothesis):

There is no significant relationship between global temperature anomalies and the concentration of CO₂ and CH₄ from 2000 to 2010.
Greenhouse gas concentrations do not have a measurable impact on global temperature variations over the studied period.
H₁ (Alternative Hypothesis):

There is a statistically significant positive correlation between CO₂ and CH₄ concentrations and global temperature anomalies.
Rising levels of these greenhouse gases contribute to an increase in temperature anomalies over time.
```{r}
summary(df_ch4)
```

```{r}
sum(is.na(df_ch4))
```

```{r}
data_ch4 = merge(df_ch4, df_anomaly, by = c("Year", "Month"), all.x = TRUE)
data_ch4 <- data_ch4[order(data_ch4$Year, data_ch4$Month), ]
print(data_ch4, row.names = FALSE)
```

## Handling Missing Values

```{r}
cat("Dimension of data:", dim(data_ch4))
cat("\nNumber of N/A values", sum(is.na(data_ch4)))
```

## Explore Data Characteristics

```{r}
num_col = c("average_ch4", "trend_ch4", "Anomaly")
data_wo_date = data_ch4[, num_col]
```

```{r}
# Summary statistics for all columns
summary(data_ch4)
```

```{r}
# Mean for each numeric column
cat("\nMean for each numeric column\n")
colMeans(data_ch4[, sapply(data_ch4, is.numeric)], na.rm = TRUE)

# Standard deviation for each numeric column
cat("\nStandard deviation for each numeric column\n")
sapply(data_ch4[, sapply(data_ch4, is.numeric)], sd, na.rm = TRUE)

# Median for each numeric column
cat("\nMedian for each numeric column\n")
sapply(data_ch4[, sapply(data_ch4, is.numeric)], median, na.rm = TRUE)

# Min and max values
cat("\nMin and Max of each numeric column\n")
sapply(data_ch4[, sapply(data_ch4, is.numeric)], range, na.rm = TRUE)
```
```{r}
cal_sum_stat = function(df){
  num_col = c("average_ch4", "trend_ch4", "Anomaly")
  data_wo_date = data_ch4[, num_col]
  
  # Apply function to each column
  stats <- sapply(data_wo_date, function(col) {
    c(Mean = mean(col, na.rm = TRUE), Median = median(col, na.rm = TRUE), Min = min(col, na.rm = TRUE),
      Max = max(col, na.rm = TRUE), Quantiles = quantile(col, na.rm = TRUE),
      Standard_Deviation = sd(col, na.rm = TRUE), Variance = var(col, na.rm = TRUE))
  })
  
  return(as.data.frame(t(stats)))
}

result = cal_sum_stat(data_ch4)
print(result)
```
ok 
```{r}
#data %>%
#  group_by(Year) %>%
#  summarise(
#    mean_anomaly = mean(Anomaly, na.rm = TRUE),
#    sd_anomaly = sd(Anomaly, na.rm = TRUE),
#    min_anomaly = min(Anomaly, na.rm = TRUE),
#    max_anomaly = max(Anomaly, na.rm = TRUE)
#  )

```

## Visualize Data Relationships

### Univariate Analysis

```{r}
data_ch4 %>%
  group_by(Month) %>%
  summarise(mean_anomaly = mean(Anomaly, na.rm=TRUE)) %>%
  ggplot(aes(x=factor(Month), y=mean_anomaly)) +
  geom_col(fill="lightblue") +
  labs(title="Average Anomaly by Month", x="Month", y="Anomaly")

```
```{r}
# List of numeric columns
num_cols <- c("average_ch4", "trend_ch4", "Anomaly")

# Loop through each column and generate a box plot
for (col in num_cols) {
  p <- ggplot(data_ch4, aes(y = .data[[col]])) +
    geom_boxplot(fill = "blue", color = "black") +
    labs(title = paste("Box Plot of", col),
         y = col) +
    theme_minimal()
  
  print(p)  # Display each plot separately
}

```
```{r}
# Loop through each column and generate a line plot
for (col in num_cols) {
  p <- ggplot(data_ch4, aes(x = Year + (Month - 1) / 12, y = .data[[col]])) +
    geom_line(color = "blue", size = 1) +
    labs(title = paste("Line Plot of", col, "Over Time"),
         x = "Year",
         y = col) +
    theme_minimal()
  
  print(p)  # Display each plot separately
}
```
```{r}
# Function to create histogram with KDE for each numerical column
plot_hist_kde <- function(df) {
  numeric_cols <- names(df)[sapply(df, is.numeric)]  # Get numeric columns
  
  for (col in numeric_cols) {
    p <- ggplot(df, aes(x = .data[[col]])) +
      geom_histogram(aes(y = ..density..), bins = 30, fill = "blue", alpha = 0.5) +  # Histogram
      geom_density(color = "red", size = 1.2) +  # KDE (Kernel Density Estimate)
      labs(title = paste("Histogram & KDE of", col), x = col, y = "Density") +
      theme_minimal()
    
    print(p)  # Print each plot separately
  }
}

# Apply function to your dataset
plot_hist_kde(data_wo_date)

```

### Bivariate Analysis

```{r}
# Create line plot
ggplot(data_ch4, aes(x = Year + (Month - 1) / 12)) +  # Convert Year and Month into a continuous time axis
  geom_line(aes(y = average_ch4, color = "Average CH4"), size = 1) +
  geom_line(aes(y = trend_ch4, color = "Trend CH4"), size = 1) +
  scale_color_manual(values = c("Average CH4" = "blue", "Trend CH4" = "red")) +
  scale_x_continuous(breaks = seq(2000, 2010, 1)) +  # Ensure all years are displayed
  labs(title = "CH4 Trend vs Average CH4 Over Time",
       x = "Year",
       y = "CH4 Levels",
       color = "Legend") +
  theme_minimal()

```
```{r}
ggplot(data_ch4, aes(x = average_ch4, y = Anomaly)) +
  geom_point(color = "blue", alpha = 0.6) +  # Scatter points
  geom_smooth(method = "lm", color = "red", se = TRUE) +  # Regression line
  labs(title = "Scatter Plot of CH4 vs Anomaly",
       x = "Average CH4 Levels",
       y = "Anomaly") +
  theme_minimal()

```
```{r}
ggplot(data_ch4, aes(x = average_ch4, y = Anomaly)) +
  geom_hex(bins = 30) +  # Hexbin plot for density visualization
  scale_fill_viridis_c() +  # Color scale for density
  labs(title = "Hexbin Density Plot of CH4 vs Anomaly",
       x = "Average CH4 Levels",
       y = "Anomaly") +
  theme_minimal()
```
```{r}
ggplot(data_ch4, aes(x = cut(average_ch4, breaks = 5), y = Anomaly)) +
  geom_boxplot(fill = "lightblue", color = "black") +
  labs(title = "Box Plot of Anomaly Across CH4 Levels",
       x = "CH4 Level Ranges",
       y = "Anomaly") +
  theme_minimal()
```
```{r}
correlation <- cor(data_ch4$average_ch4, data_ch4$Anomaly, use = "complete.obs")
print(correlation)  # Prints correlation coefficient
```

### Multivariate Analysis

```{r}
# creating correlation matrix
corr_mat <- round(cor(data_wo_date),2)
 
# reduce the size of correlation matrix
melted_corr_mat <- melt(corr_mat)
head(melted_corr_mat)
 
# plotting the correlation heatmap
ggplot(data = melted_corr_mat, aes(x=Var1, y=Var2, 
                                   fill=value)) + 
  labs(title = "Correlation Heatmap") +
  geom_tile() +
  geom_text(aes(Var2, Var1, label = value), 
          color = "white", size = 8)
```
```{r}
boxplot_stats <- function(df) {
  numeric_cols <- names(df)[sapply(df, is.numeric)]
  
  for (col in numeric_cols) {
    outliers <- boxplot.stats(df[[col]])$out
    print(paste("Outliers in", col, ":", paste(outliers, collapse = ", ")))
  }
}

boxplot_stats(data_ch4)  # Replace with your dataframe name
```

```{r}
remove_outliers <- function(df) {
  numeric_cols <- names(df)[sapply(df, is.numeric)]
  
  for (col in numeric_cols) {
    Q1 <- quantile(df[[col]], 0.25, na.rm = TRUE)
    Q3 <- quantile(df[[col]], 0.75, na.rm = TRUE)
    IQR <- Q3 - Q1
    lower_bound <- Q1 - 1.5 * IQR
    upper_bound <- Q3 + 1.5 * IQR
    df <- df[df[[col]] >= lower_bound & df[[col]] <= upper_bound, ]
  }
  
  return(df)
}

data_ch4_clean <- remove_outliers(data_ch4)  # New dataframe without outliers

```

```{r}
winsorize <- function(df) {
  numeric_cols <- names(df)[sapply(df, is.numeric)]
  
  for (col in numeric_cols) {
    Q1 <- quantile(df[[col]], 0.25, na.rm = TRUE)
    Q3 <- quantile(df[[col]], 0.75, na.rm = TRUE)
    IQR <- Q3 - Q1
    lower_bound <- Q1 - 1.5 * IQR
    upper_bound <- Q3 + 1.5 * IQR
    
    df[[col]] <- ifelse(df[[col]] < lower_bound, lower_bound,
                        ifelse(df[[col]] > upper_bound, upper_bound, df[[col]]))
  }
  
  return(df)
}

data_ch4_winsorized <- winsorize(data_ch4)  # Capped outliers

```

```{r}
# Load required library
library(ggplot2)

# Function to plot boxplots for each numeric column
plot_boxplots <- function(df, title_suffix = "") {
  numeric_cols <- names(df)[sapply(df, is.numeric)]
  
  for (col in numeric_cols) {
    print(
      ggplot(df, aes(y = .data[[col]])) +
        geom_boxplot(fill = "lightblue", color = "black", outlier.color = "red", outlier.shape = 16) +
        labs(title = paste("Boxplot of", col, title_suffix),
             y = col) +
        theme_minimal()
    )
  }
}

# Boxplots after removing outliers
plot_boxplots(data_ch4_clean, "(After Removing Outliers)")

# Boxplots after Winsorization
plot_boxplots(data_ch4_winsorized, "(After Winsorization)")

```

